Evaluating Covid-19 Short-Term Forecasts using \pkg{scoringutils} in \proglang{R}

### 2021-12-10 article-knitr.Rnw
switch to knitr, update a few plots

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE, prefix.string=plots/plot}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)

@@

library("MASS")
library(patchwork)
library(ggplot2)
library(kableExtra)
library(scoringutils)
library(data.table)
library(kableExtra)
library(magrittr)
library(knitr)

@@

\begin{document}
\SweaveOpts{concordance=FALSE}

@@

\begin{table}[h!]
\centering
<<forecast-types-table, echo=FALSE, results=tex>>=

@@

<<summary-table, echo=FALSE, eval = FALSE, results=tex>>=

@@

\newpage
\begin{table}[h!]
\centering
<<summary-table, echo=FALSE, results=tex>>=

@@

\begin{figure}[h]
\centering
<<example-calibration, echo=FALSE, fig=TRUE, height=7.2, width=14, eval = FALSE>>=
n_truth = 1000
n_samples = 100
true_values <- rnorm(n_truth, 0, 1)
predictions1 <- rnorm(n_truth * n_samples)
predictions2 <- rnorm(n_truth * n_samples, mean = 0.5)
predictions3 <- rnorm(n_truth * n_samples, sd = 1.4)
predictions4 <- rnorm(n_truth * n_samples, sd =  0.7)
# predictions5 <- rnorm(n_truth * n_samples, mean = 0.2, sd =  0.7)

df <- data.table::data.table(true_value = rep(true_values, each = n_samples), 
                             id = rep(1:n_truth, each = n_samples),
                             prediction = c(predictions1, predictions2, 
                                            predictions3, predictions4), 
                             sample = 1:n_samples,
                             model = rep(c("Normal(0, 1)", "Normal(0.5, 1)", 
                                       "Normal(0, 1.4)", "Normal(0, 0.7)"), 
                                       each = n_truth * n_samples))

res <- eval_forecasts(df, 
                      by = c("model", "id"), 
                      summarise_by = c("model"), 
                      pit_plots = TRUE)

# plot with observations
true_value_plot <- ggplot2::ggplot(data = data.frame(x = true_values),
                ggplot2::aes(x = x)) +
  ggplot2::geom_histogram(ggplot2::aes(y = ..density..),
                          fill = "grey",
                          colour = "dark grey") +
  cowplot::theme_cowplot() +
  ggplot2::labs(x = "True values",
                y = "Density") +
  ggplot2::theme(legend.position = "bottom")

# plot with standard normal distribution
standard_normal <- true_value_plot +
  ggplot2::geom_function(fun = dnorm, colour = "black") +
  ggplot2::ggtitle("Normal(0, 1)")

# plot with shifted mean
shifted_mean <- true_value_plot +
  ggplot2::geom_function(fun = dnorm, colour = "black", args = list(mean = 0.5)) +
  ggplot2::ggtitle("Normal(0.5, 1)")

# plot with overdispersion
overdispersion <- true_value_plot +
  ggplot2::geom_function(fun = dnorm, colour = "black", args = list(sd = 1.4)) +
  ggplot2::ggtitle("Normal(0, 1.4)")

# plot with underdispersion
underdispersion <- true_value_plot +
  ggplot2::geom_function(fun = dnorm, colour = "black", args = list(sd = 0.7)) +
  ggplot2::ggtitle("Normal(0, 0.7)")

scores_table <- dcast(melt(res$scores, id.vars = "model", 
                           variable.name = "score"), 
                      score ~ model)
setcolorder(
  scores_table, 
  c("score", "Normal(0, 1)", "Normal(0.5, 1)", "Normal(0, 1.4)", "Normal(0, 0.7)")
)

(standard_normal | shifted_mean | overdispersion | underdispersion) /
  (res$pit_plots$`Normal(0, 1)` | res$pit_plots$`Normal(0.5, 1)` | res$pit_plots$`Normal(0, 1.4)` | res$pit_plots$`Normal(0, 0.7)`) / 
  gridExtra::tableGrob(scores_table)

ggsave("plots/calibration-diagnostic-examples.png", width = 12.5, height = 6)
@

@@

combined <- merge_pred_and_obs(uk_forecasts, covid_uk_data)
@
% 
It is helpful to start the evaluation process by visualising the available data, as missing forecasts can impact the evaluation if missingness is not random, but instead correlates with performance. The function \fct{show\_avail\_forecasts} returns a heatmap with the number of available forecasts. By default, the function treats a set of different quantiles or samples as one forecast. However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts. 
% 
\begin{figure}[h]
\centering
<<show-availability, echo=TRUE, fig=TRUE, height=6.7, width=12, eval = TRUE>>=
show_avail_forecasts(data = combined, 
                     x = "creation_date", 
                     show_numbers = FALSE,
                     legend_position = "bottom",
                     facet_formula = ~ value_desc)
@
\caption{\label{fig:avail-forecasts} Overview of the number of available forecasts}
\end{figure}
%THINK ABOUT WHETHER THIS FUNCTION CAN BE COMBINED WITH THE OTHER HEATMAP FUNCTION. 
% 
The forecasts and observed values themselves can be visualised using the \fct{plot\_predictions} function. Data visualisation is of course highly context-dependent, but the function tries to accommodate most practical use-cases. In addition to basic plotting functionality it offers the user an optional ad hoc way to filter both forecasts and observed values. This makes it possible to tweak the plot in a beginner-friendly way without having to manipulate the data separately. Forecasts and observed values can be passed in separately (and are merged internally) or as a single data.frame. Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data. To display, for example, death forecasts from the SIRCOVID model made on the 22nd July 2020 for the four UK countries and add three weeks of prior observations to the plot, we can call: % 
\begin{figure}[h!]
\centering
<<show-forecasts, echo=TRUE, fig=TRUE, height=5.2, width=7, eval = FALSE>>=
locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
plot_predictions(truth = covid_uk_data,
                 forecasts = uk_forecasts,
                 filter_both = list(paste("geography %in%", locations, collapse = ""), 
                                    'value_type == "death_inc_line"'),
                 filter_truth = list('value_date <= "2020-06-22"',
                                     'value_date > "2020-06-01"'), 
                 filter_forecasts = list('model == "SIRCOVID"', 
                                         'creation_date == "2020-06-22"'),
                 x = "value_date",
                 facet_formula = ~ geography) + 
  ggplot2::theme(legend.position = "bottom")
@
\caption{\label{fig:forecast-visualisation} Short-term forecasts for incident deaths made by the SIRCOVID model on June 22 2020.}
\end{figure}
% 
The output is shown in Figure \ref{fig:forecast-visualisation}.

@@

\begin{figure}[h]
\centering
<<score-table, echo=TRUE, fig=TRUE, height=7.2, width=15, eval = TRUE>>=
score_table(scores, y = "model", facet_formula = ~ value_desc)
@
\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
\end{figure}

@@

\begin{figure}[h!]
\centering
<<pairwise-plot, echo=TRUE, fig=TRUE, height=7.2, width=14, eval = TRUE>>=
plot_pairwise_comparison(pairwise) +
  ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
@
\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets. If a tile is blue, then the model on the y-axis performed better. If it is red, the model on the x-axis performed better in direct comparison. }
\end{figure}

@@

\begin{figure}[h!]
\centering
<<WIS-components, echo=TRUE, fig=TRUE, height=5.2, width=12, eval = TRUE>>=
wis_components(scores, 
               facet_formula = ~ value_desc, 
               scales = "free_x", 
               relative_contributions = TRUE, 
               x_text_angle = 0) + 
  ggplot2::coord_flip() + 
  ggplot2::theme(legend.position = "bottom")
@
\caption{\label{fig:wis-components} p-values and ratios together.}
\end{figure}

@@

\begin{figure}[h!]
\centering
<<coverage, echo=TRUE, fig=TRUE, height=5.2, width=14, out.width = '0.4\\linewidth', fig.show='hold', eval = TRUE>>=
cov_scores <- eval_forecasts(combined, 
                             summarise_by = c("model", 
                                              "range", "quantile"))
scoringutils::interval_coverage(cov_scores)
scoringutils::quantile_coverage(cov_scores)
@

@@

\begin{figure}[h]
\centering
<<calibration, echo=TRUE, fig=TRUE, height=5.2, width=15, eval = TRUE>>=
scores <- eval_forecasts(combined, 
                         summarise_by = c("model",
                                          "value_desc", 
                                          "geography"), 
                         compute_relative_skill = FALSE)
scoringutils::score_heatmap(scores, metric = "bias", 
                            x = "geography", facet_formula = ~ value_desc, 
                            scale = "free_x")
@
\caption{\label{fig:bias-heatmap} bias by location}
\end{figure}

@@

\begin{table}[h!]
\centering
<<scores-table, echo=FALSE, results=tex>>=

>>>>

%% For Sweave-based articles about R packages:
%% need no \usepackage{Sweave}

<<include=FALSE>>=
library(knitr)
opts_chunk$set(
  engine='R', 
  tidy=TRUE, 
  fig.path='plots/plot-', 
  message = FALSE, 
  warning = FALSE
)
@

<<preliminaries, echo=FALSE, results='hide'>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)

@@

library("MASS")
library(patchwork)
library(ggplot2)
library(kableExtra)
library(scoringutils)
library(data.table)
library(kableExtra)
library(magrittr)
library(knitr)
library(here)

@@

\begin{document}

<<include=FALSE>>=
library(knitr)
opts_chunk$set(
concordance=FALSE
)
@

@@

\begin{table}[h!]
\centering
<<forecast-types-table, echo=FALSE, results='asis'>>=

@@

<<summary-table, echo=FALSE, eval = FALSE, results='asis'>>=

@@

\newpage
\begin{table}[h!]
\centering
<<summary-table2, echo=FALSE, results='asis'>>=

@@

\begin{figure}[h]
\centering
<<example-calibration2, echo=FALSE, fig.height=7.2, fig.width=14, eval = TRUE>>=
knitr::include_graphics(here("plots", "calibration-diagnostic-examples.png"))
@

@@

combined <- merge_pred_and_obs(uk_forecasts, covid_uk_data)
@
% 
It is helpful to start the evaluation process by visualising the available data, as missing forecasts can impact the evaluation if missingness is not random, but instead correlates with performance. The function \fct{show\_avail\_forecasts} returns a heatmap with the number of available forecasts. By default, the function treats a set of different quantiles or samples as one forecast. However, the user can specify manually which elements to treat as one forecast and which categories to sum over to count the number of available forecasts. 
% 
\begin{figure}[h]
\centering
<<show-availability, echo=TRUE, fig.height=6.7, fig.width=12, eval = TRUE>>=
show_avail_forecasts(data = combined, 
                     x = "creation_date", 
                     show_numbers = FALSE,
                     legend_position = "bottom",
                     facet_formula = ~ value_desc)
@
\caption{\label{fig:avail-forecasts} Overview of the number of available forecasts}
\end{figure}
%THINK ABOUT WHETHER THIS FUNCTION CAN BE COMBINED WITH THE OTHER HEATMAP FUNCTION. 
% 
The forecasts and observed values themselves can be visualised using the \fct{plot\_predictions} function. Data visualisation is of course highly context-dependent, but the function tries to accommodate most practical use-cases. In addition to basic plotting functionality it offers the user an optional ad hoc way to filter both forecasts and observed values. This makes it possible to tweak the plot in a beginner-friendly way without having to manipulate the data separately. Forecasts and observed values can be passed in separately (and are merged internally) or as a single data.frame. Conditions to filter on need to be provided as a list of strings, where each of the strings represents an expression that can be evaluated to filter the data. To display, for example, death forecasts from the SIRCOVID model made on the 22nd July 2020 for the four UK countries and add three weeks of prior observations to the plot, we can call: % 
\begin{figure}[h!]
\centering
<<show-forecasts, echo=TRUE, fig.height=5.2, fig.width=7, eval = FALSE>>=
locations <- 'c("Scotland", "Wales", "Northern Ireland", "England")'
plot_predictions(truth = covid_uk_data,
                 forecasts = uk_forecasts,
                 filter_both = list(paste("geography %in%", locations, collapse = ""), 
                                    'value_type == "death_inc_line"'),
                 filter_truth = list('value_date <= "2020-06-22"',
                                     'value_date > "2020-06-01"'), 
                 filter_forecasts = list('model == "SIRCOVID"', 
                                         'creation_date == "2020-06-22"'),
                 x = "value_date",
                 facet_formula = ~ geography) + 
  ggplot2::theme(legend.position = "bottom")
@
\caption{\label{fig:forecast-visualisation} Short-term forecasts for incident deaths made by the SIRCOVID model on June 22 2020.}
\end{figure}
% 
The output is shown in Figure \ref{fig:forecast-visualisation}.

@@

\begin{figure}[h]
\centering
<<score-table, echo=TRUE, fig.height=7.2, fig.width=15, eval = TRUE>>=
score_table(scores, y = "model", facet_formula = ~ value_desc)
@
\caption{\label{fig:score-table} Coloured table to visualise the computed scores}
\end{figure}

@@

\begin{figure}[h!]
\centering
<<pairwise-plot, echo=TRUE, fig.height=7.2, fig.width=14, eval = TRUE>>=
plot_pairwise_comparison(pairwise) +
  ggplot2::facet_wrap(~ value_desc, scales = "free_x") 
@
\caption{\label{fig:pairwise-comparison} Ratios of mean scores based on overlapping forecast sets. If a tile is blue, then the model on the y-axis performed better. If it is red, the model on the x-axis performed better in direct comparison. }
\end{figure}

@@

\begin{figure}[h!]
\centering
<<WIS-components, echo=TRUE, fig.height=5.2, fig.width=12, eval = TRUE>>=
wis_components(scores, 
               facet_formula = ~ value_desc, 
               scales = "free_x", 
               relative_contributions = TRUE, 
               x_text_angle = 0) + 
  ggplot2::coord_flip() + 
  ggplot2::theme(legend.position = "bottom")
@
\caption{\label{fig:wis-components} p-values and ratios together.}
\end{figure}

@@

\begin{figure}[h!]
\centering
<<coverage, echo=TRUE, fig.height=5.2, fig.width=14, out.width="40%", fig.show='hold', eval = TRUE>>=
cov_scores <- eval_forecasts(combined, 
                             summarise_by = c("model", 
                                              "range", "quantile"))
scoringutils::interval_coverage(cov_scores)
scoringutils::quantile_coverage(cov_scores)
@

@@

\begin{figure}[h]
\centering
<<calibration, echo=TRUE, fig.height=5.2, fig.width=15, eval = TRUE>>=
scores <- eval_forecasts(combined, 
                         summarise_by = c("model",
                                          "value_desc", 
                                          "geography"), 
                         compute_relative_skill = FALSE)
scoringutils::score_heatmap(scores, metric = "bias", 
                            x = "geography", facet_formula = ~ value_desc, 
                            scale = "free_x")
@
\caption{\label{fig:bias-heatmap} bias by location}
\end{figure}

@@

\begin{table}[h!]
\centering
<<scores-table, echo=FALSE, results='asis'>>=
